---
title: "Invasiveness and Rarity across Angiosperm Families"
author: "Maxwell J. Farrell"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
  html_document:
    highlight: default
    toc: no
    toc_depth: 2
    toc_float: true
    theme: yeti
urlcolor: blue
---

<!-- knit with Rscript -e "rmarkdown::render('angiosperm_invasion.Rmd')" -->

## JD Notes

Use ratio of threatened to non-threatened species listed by IUCN (to account for reporting biases by IUCN)

Make sure we aren't losing families because we are picking a representative species that isn't there...

Calculate clade age before pruning (using whole zanne tree)

Use plant list to get representatives from all families before calculating  clade age

"Rare" is actually "threatened" based on IUCN

Look at the relationships between woodiness and climate sum with family size (larger families, mroe chance to have variable life histories...)


JD cape flora & mark schwartz (california) showed positive relationship with rarity and clade size (centers of rapid diversification and endemism), but not sure it extends to global. Nobody *maybe* has shown opposite relationship with invasiveness


"Expanding and declining species" 

## Questions

1. Are plant families with more rare species more uniform in functional traits?
2. Are plant families with more invasive species more heterogeneous in functional traits?
3. Are rarity and invasiveness correlated across families?

Outcomes to show:

- Sampling effect of clade size
- Differences from predicted values
- Family effects



## Parasite / pathogen impacts

Get grass fungus - host associations 
Darwin's naturalization hypothesis & deviations as a result of specialism / generalism of their known pathogens



## Introduction Outline

- Traits underlying rarity and invasiveness have been studied before and are important for XXX
- These patterns have (rarely, or never?) been analyzed at the global scale 
- The potential for rarity and invasiveness may increase with number of species in a clade (pure sampling effect), leading to the prediction that both rarea and invasive species will be found in larger families (rarity and invasiveness are positively correlated). 
- However, different functional traits are likely to be underlying rarity and invasiveness (examples)
- Here we present a family level comparative analyses of rarity and invasiveness for all plants, and identify XX and YYY 


- Family effects

## To Do

- Try models without family age (is this redundant with a phylogenetic covariance matrix?)

- Try without family effect (to see if this causes the directionaliry shift from JPs models)

- Try without Orchids (sensitivity analysis)

- Potential other predictors: number of known hybrids in a family

- Posterior predictions - which families have more invasive, or less rare species than predicted from the model? FOR DISCUSSION

- correlation (regression) of rarity versus invasive (with size as only other predictor)


##Packages
```{r loading packages, echo=T, message=F, warning=F}
require(tidyr)
require(dplyr)
require(magrittr)
require(ggplot2)
require(bayesplot)
require(ape)
require(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
require(brms)
require(ggmcmc)
require(knitr)
require(phytools)
# require(Taxonstand)

```

##Loading data

```{r loading data, echo=T, message=F, warning=F, results='hide'}

data <- read.csv("./sp.rare.inv.csv", as.is=F)
# data %>% purrr::keep(is.numeric) %>% cor(., use="complete.obs") %>% View()

# one family has more rare species than species
data[data$rare>data$species,]

# increasing species by 1 to match
data$species[data$family=="Metteniusaceae"] <- 8

# creating column for woodiness
data$woodiness <- data$herbaceous + data$woody + data$variable

### Zanne tree & generating tree representing families
tree <- read.tree("./Vascular_Plants_rooted.dated.tre")
spec_fam <- read.csv("./spec.fam.csv", as.is=T)
tree <- drop.tip(tree, setdiff(tree$tip.label,as.character(spec_fam$sp)))
lookup <- setNames(spec_fam$family, spec_fam$sp)
tree$tip.label <- as.character(lookup[tree$tip.label])

# Getting new family age
min_above_zero <- function(x){

  if (length(x)>1) {
    min(x[x>0])
  } else {0}

}

# Calculate family age (may be influenced by families missing from the tree)
coph <- cophenetic(tree)

fam_age <- sapply(rownames(coph), function(fam) 
          min_above_zero(coph[rownames(coph)%in%fam,])
          ) 
old_ages <- data %>% select(family, age)
new_ages <- data.frame(family=names(fam_age), new_age=as.numeric(fam_age)/2)
new_ages %<>% arrange(family)
data <- left_join(data,new_ages)

# with(data, plot(age ~ new_age, pch=20))
# with(data, cor(age,new_age))

# creating another family level variable (for non-phylogenetic family level effects)
data$family_name <- data$family

# These families we have data for, but are not in tree
setdiff(data$family, tree$tip.label)
# intersect(data$family, tree$tip.label)

# remove missing families in data
data <- data[data$family%in%tree$tip.label,]

### Davies tree
# tree <- read.tree("./ultrametric.phy")

# # some families missing...
# length(intersect(data$family, tree$tip.label))
# setdiff(data$family, tree$tip.label)
# setdiff(tree$tip.label, data$family)

# # drop these from the tree
# tree <- drop.tip(tree, setdiff(tree$tip.label,as.character(data$family)))

# # remove missing families in data
# data <- data[data$family%in%tree$tip.label,]

# phylogenetic correlation structure
phy_cov <- vcv(tree, corr=TRUE)

```
&nbsp;
&nbsp;

<!-- ##Histograms of numeric predictors -->

```{r histograms, echo=F, eval=F}
ggplot(gather(data[,2:21]%>% purrr::keep(is.numeric)), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free')

ggplot(gather(data[,22:41]%>% purrr::keep(is.numeric)), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free')

ggplot(gather(data[,42:65]%>% purrr::keep(is.numeric)), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free')


```

<!-- ## Pairs of predictors-->

```{r pred_pairs, echo=F, eval=F}

dat2 <- data[,c("forms2","sum.photo","breeding.sum","climate.sum","age","species")]
dat2$age <- log10(dat2$age)
dat2$species <- log10(dat2$species)

pairs(dat2)

with(data, plot(breeding.sum, breeding.sum2))
with(data, cor(breeding.sum, breeding.sum2))

with(data, plot(forms, forms2))
with(data, cor(forms, forms2))

with(data, plot(sum.photo, sum.photo2))
with(data, cor(sum.photo, sum.photo2))

```

## Relationship between rarity and invasiveness categories

```{r rare_invasive, echo=TRUE, eval=TRUE}

if (!file.exists("./rare_invasive_2k.rds")) {

  rare_inv <- brm(rare | trials(vetted) ~ log10(Invasive+1) + log10(new_age) +log10(species) +  (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin = 4, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.95,max_treedepth=12), cores=4)

  saveRDS(rare_inv, "./rare_invasive_2k.rds")

} else { rare_inv <- readRDS("./rare_invasive_2k.rds")}

# var_family <- posterior_samples(rare_inv, pars = "sd_family__Intercept")
# var_family_name <- posterior_samples(rare_inv, pars = "sd_family_name__Intercept")
# varProp_full <- (var_family^2)/(var_family^2 + var_family_name^2)
# median(varProp_full)
# plot(density(varProp_full))

# var_familyFULL <- (var_family^2+var_family_name^2)/(var_family^2 + var_family_name^2 + pi*2/3)
# plot(density(var_familyFULL))

hyp <- paste("sd_family__Intercept^2 /", "(sd_family__Intercept^2 + sd_family_name__Intercept^2) = 0")
(hyp <- hypothesis(rare_inv, hyp, class = NULL))
plot(hyp)

# ########################

# rare_inv

outcome <- summary(rare_inv$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
to_print <- to_print[-1,]

names1 <- c("Number of invasive species","Family age","Number of species","sd phylogenetic effects","sd family specific effects")

rare_inv_plot <- stanplot(rare_inv, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean") + scale_y_discrete(labels=rev(names1), limits=rev(rownames(to_print)))

rownames(to_print) <- names1
kable(to_print, digits=2)

rare_inv_plot

bayes_R2(rare_inv)

##Posterior predictive checks
# first make posterior predictions
pp_rare_inv <- brms::posterior_predict(rare_inv)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=rare_inv$data$rare, pp_rare_inv[1:50, ]) + scale_x_continuous(trans="log1p")


if (!file.exists("./rare_naturalized_2k.rds")) {

  rare_nat <- brm(rare | trials(vetted) ~ log10(Naturalised+1) + log10(new_age) +log10(species) +  (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.95,max_treedepth=12), cores=4)

  saveRDS(rare_nat, "./rare_naturalized_2k.rds")

} else { rare_nat <- readRDS("./rare_naturalized_2k.rds")}

# rare_nat

outcome <- summary(rare_nat$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
to_print <- to_print[-1,]

names1 <- c("Number of naturalized species","Family age","Number of species","sd phylogenetic effects","sd family specific effects")

rare_nat_plot <- stanplot(rare_nat, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean") + scale_y_discrete(labels=rev(names1), limits=rev(rownames(to_print)))

rownames(to_print) <- names1
kable(to_print, digits=2)

rare_nat_plot

bayes_R2(rare_nat)

##Posterior predictive checks
# first make posterior predictions
pp_rare_nat <- brms::posterior_predict(rare_nat)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=rare_nat$data$rare, pp_rare_nat[1:50, ]) + scale_x_continuous(trans="log1p")


if (!file.exists("./rare_environmental_2k.rds")) {

  rare_env <- brm(rare | trials(vetted) ~ log10(Environ+1) + log10(new_age) +log10(species) +  (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.95,max_treedepth=12), cores=4)

  saveRDS(rare_env, "./rare_environmental_2k.rds")

} else { rare_env <- readRDS("./rare_environmental_2k.rds")}

# rare_env

outcome <- summary(rare_env$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
to_print <- to_print[-1,]

names1 <- c("Number of environmental species","Family age","Number of species","sd phylogenetic effects","sd family specific effects")

rare_env_plot <- stanplot(rare_env, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean") + scale_y_discrete(labels=rev(names1), limits=rev(rownames(to_print))) 

rownames(to_print) <- names1
kable(to_print, digits=2)

rare_env_plot

bayes_R2(rare_env)

##Posterior predictive checks
# first make posterior predictions
pp_rare_env <- brms::posterior_predict(rare_env)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=rare_env$data$rare, pp_rare_env[1:50, ]) + scale_x_continuous(trans="log1p")

```

```{r, echo=F, eval=F}
## Forest plots

require(gridExtra)
grid.arrange(rare_inv_plot,rare_nat_plot,rare_env_plot, ncol=3)

```


## Proportion rare species
```{r frac_rare, echo=TRUE, eval=TRUE}

if (!file.exists("./fit_rare_2k.rds")) {

  fit_rare <- brm(rare | trials(vetted) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.95,max_treedepth=13))

  saveRDS(fit_rare, "./fit_rare_2k.rds")

} else { fit_rare <- readRDS("./fit_rare_2k.rds")}

# fit_rare

outcome <- summary(fit_rare$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
to_print <- to_print[-1,]

names1 <- c("Woodiness","Climate sum","Number of species hybrids","Family age","Number of species","Hermaphroditic","Dioeceous","Monoeceous","Asexual","sd phylogenetic effects","sd family specific effects")

fit_rare_plot <- stanplot(fit_rare, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean") + scale_y_discrete(labels=rev(names1), limits=rev(rownames(to_print))) + ggtitle("Number of rare species")

rownames(to_print) <- names1
kable(to_print, digits=2)

fit_rare_plot

bayes_R2(fit_rare)

##Posterior predictive checks
# first make posterior predictions
pp_rare <- brms::posterior_predict(fit_rare)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_rare$data$rare, pp_rare[1:50, ]) + scale_x_continuous(trans="log1p")


```


## Proportion naturalized species
```{r frac_nat, echo=TRUE, eval=TRUE}

if (!file.exists("./fit_nat_2k.rds")) {

  fit_nat <- brm(Naturalised | trials(species) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.8, max_treedepth=11))

  saveRDS(fit_nat, "./fit_nat_2k.rds")

} else { fit_nat <- readRDS("./fit_nat_2k.rds")}

# fit_nat

outcome <- summary(fit_nat$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
to_print <- to_print[-1,]

names1 <- c("Woodiness","Climate sum","Number of species hybrids","Family age","Number of species","Hermaphroditic","Dioeceous","Monoeceous","Asexual","sd phylogenetic effects","sd family specific effects")

fit_nat_plot <- stanplot(fit_nat, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean") + scale_y_discrete(labels=rev(names1), limits=rev(rownames(to_print)))+ ggtitle("Number of naturalized species")

rownames(to_print) <- names1
kable(to_print, digits=2)

fit_nat_plot

bayes_R2(fit_nat)

##Posterior predictive checks
# first make posterior predictions
pp_nat <- brms::posterior_predict(fit_nat)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_nat$data$Naturalised, pp_nat[1:50, ]) + scale_x_continuous(trans="log1p")


```


## Proportion invasive species
```{r frac_inv, echo=TRUE, eval=TRUE}

# Invasive

if (!file.exists("./fit_inv_2k.rds")) {

  fit_inv <- brm(Invasive | trials(species) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.8,max_treedepth=12))

  saveRDS(fit_inv, "./fit_inv_2k.rds")

} else { fit_inv <- readRDS("./fit_inv_2k.rds")}

# fit_inv

outcome <- summary(fit_inv$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
to_print <- to_print[-1,]

names1 <- c("Woodiness","Climate sum","Number of species hybrids","Family age","Number of species","Hermaphroditic","Dioeceous","Monoeceous","Asexual","sd phylogenetic effects","sd family specific effects")

fit_inv_plot <- stanplot(fit_inv, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean") + scale_y_discrete(labels=rev(names1), limits=rev(rownames(to_print))) + ggtitle("Number of invasive species")

rownames(to_print) <- names1
kable(to_print, digits=2)

fit_inv_plot

bayes_R2(fit_inv)

##Posterior predictive checks
# first make posterior predictions
pp_inv <- brms::posterior_predict(fit_inv)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_inv$data$Invasive, pp_inv[1:50, ]) + scale_x_continuous(trans="log1p")

```


```{r, echo=F, eval=F}
# Forest plots

require(gridExtra)
grid.arrange(fit_rare_plot,fit_nat_plot,fit_inv_plot, ncol=3)

```


## Phylogenetic effects plots

```{r, fig.width=8, fig.height=8}

# Extract species level effects
ranef_phy <- ranef(fit_rare)$family
ranef_binom <- ranef(fit_rare)$family_name
ranef_spec <- ranef_phy + ranef_binom

ranef_spec <- as.data.frame(unlist(ranef_spec))

# re-order efects to match tree
tree_small <- drop.tip(tree, setdiff(tree$tip.label, rownames(ranef_spec)))
ranef_spec <- ranef_spec[tree_small$tip.label,]

# Plot the beside the phylogeny
cols <- c("#7570b3","#d95f02")
plot(tree_small,type="f",TRUE, cex=1,no.margin=FALSE, show.tip.label = FALSE, label.offset=10)
tiplabels(pch=21,bg=cols[ifelse(ranef_spec$Estimate.Intercept>0,1,2)],col="black",cex=abs(ranef_spec$Estimate.Intercept),adj=1.5)
legend("topright",legend=c("2","1","0","-1","-2"),pch=21,
       pt.bg=cols[c(1,1,1,2,2)],bty="n",
       text.col="gray32",cex=1, pt.cex=c(2,1,0.1,1,2))


resp <- ranef_spec$Estimate.Intercept
names(resp) <- rownames(ranef_spec)

# plotTree.barplot(tree_small, x = fastBM(tree_small), tip.labels=FALSE)
# plotTree.barplot(tree_small, x = fastBM(tree_small), args.plotTree=list(showTipLabel = FALSE))

cols2 <- NULL
cols2[resp>0] <- cols[1]
cols2[resp<0] <- cols[2]

```{r, fig.width=8, fig.height=24}

plotTree.barplot(tree_small, x = resp, args.barplot=list(xlab="Family effect",col=cols2, mar=c(5.1,0,2.1,0.6), border=NA), tip.labels=FALSE)


```




<!-- ## Proportion weed species -->

```{r frac_weed, echo=FALSE, eval=FALSE}

if (!file.exists("./fit_weed_2k.rds")) {

  fit_weed <- brm(Weed | trials(species) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.8, max_treedepth=12), cores=4)

  saveRDS(fit_weed, "./fit_weed_2k.rds")

} else { fit_weed <- readRDS("./fit_weed_2k.rds")}

fit_weed

# pairs(fit_weed)

outcome <- summary(fit_weed$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
kable(to_print, digits=2)

stanplot(fit_weed, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean")

bayes_R2(fit_weed)

##Posterior predictive checks
# first make posterior predictions
pp_weed <- brms::posterior_predict(fit_weed)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_weed$data$Weed, pp_weed[1:50, ]) + scale_x_continuous(trans="log1p")


```


<!-- ## Proportion agricultural weed species -->

```{r frac_ag, echo=FALSE, eval=FALSE}

# Agricultural.Weed

if (!file.exists("./fit_ag_2k.rds")) {

  fit_ag <- brm(Agricultural.Weed | trials(species) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.8, max_treedepth=12), cores=4)

  saveRDS(fit_ag, "./fit_ag_2k.rds")

} else { fit_ag <- readRDS("./fit_ag_2k.rds")}

fit_ag

# pairs(fit_ag)

outcome <- summary(fit_ag$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
kable(to_print, digits=2)

stanplot(fit_ag, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean")

bayes_R2(fit_ag)

##Posterior predictive checks
# first make posterior predictions
pp_ag <- brms::posterior_predict(fit_ag)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_ag$data$Agricultural.Weed, pp_ag[1:50, ]) + scale_x_continuous(trans="log1p")

```

<!-- ## Proportion Environ -->

```{r frac_env, echo=FALSE, eval=FALSE}

# Environ

if (!file.exists("./fit_env_2k.rds")) {

  fit_env <- brm(Environ | trials(species) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.8, max_treedepth=12))

  saveRDS(fit_env, "./fit_env_2k.rds")

} else { fit_env <- readRDS("./fit_env_2k.rds")}

fit_env

# pairs(fit_env)

outcome <- summary(fit_env$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
kable(to_print, digits=2)

stanplot(fit_env, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean")

bayes_R2(fit_env)

##Posterior predictive checks
# first make posterior predictions
pp_env <- brms::posterior_predict(fit_env)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_env$data$Environ, pp_env[1:50, ]) + scale_x_continuous(trans="log1p")

```

<!-- ## Proportion noxious -->

```{r frac_nox, echo=FALSE, eval=FALSE}

# Noxious

if (!file.exists("./fit_nox_2k.rds")) {

  fit_nox <- brm(Noxious | trials(species) ~ woodiness + climate.sum + shybrids + log10(new_age) + log10(species) + hermaphroditic + dioec2 + monoec2 + asexual + (1|family) + (1|family_name), 
    dat=data, family=binomial(), 
    iter=4000, thin=2, cov_ranef = list(family = phy_cov),
    control=list(adapt_delta=0.8, max_treedepth=12))

  saveRDS(fit_nox, "./fit_nox_2k.rds")

} else { fit_nox <- readRDS("./fit_nox_2k.rds")}

fit_nox

# pairs(fit_nox)

outcome <- summary(fit_nox$fit)
to_print <- outcome$summary[grep("^[sd,b]_*",rownames(outcome$summary)),]
kable(to_print, digits=2)

stanplot(fit_nox, pars=rownames(to_print), type="intervals", 
                    prob=0.8, prob_outer=0.95, point_est="mean")

bayes_R2(fit_nox)

##Posterior predictive checks
# first make posterior predictions
pp_nox <- brms::posterior_predict(fit_nox)

log1 <- scale_x_continuous(trans="log1p")

ppc_dens_overlay(y=fit_nox$data$Noxious, pp_nox[1:50, ]) + scale_x_continuous(trans="log1p")

```

